---
title: Influence of temperature and upwelling intensity on oil sardine (*Sardinella longiceps*) landings along the southwest coast of India
author: ''
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  word_document: default
  pdf_document: default
  html_document: default
header-includes:
- \usepackage[labelformat=empty]{caption}
- \usepackage{setspace}
- \onehalfspacing
- \usepackage{etoolbox}
- \AtBeginEnvironment{tabular}{\singlespacing}
- \setlength\parindent{24pt}
- \usepackage{mathptmx}
- \usepackage[margin=1in]{geometry}
- \usepackage[fontsize=12.5pt]{scrextend}
csl: '`r system.file("docs", "ecology.csl", package = "SardineForecast")`'
bibliography: '`r system.file("docs", "Sardine.bib", package = "SardineForecast")`'
---

```{r child = 'setup.Rmd'}
#Generic setup code common across all Rmds
```

# Introduction

<!-- indian oil sardine is important -->
The Indian oil sardine (*Sardinella longiceps* Valenciennes, 1847) is one of the most commercially important fish resources along the southwest coast of India (Figure `r \ref{fig:studyarea}`) and comprises ca 25% of the catch biomass [@Vivekanandanetal2003].  Landings of the Indian oilsardine are highly seasonal and peak during and after the summer monsoon period (June through September), in conjunction with the onset and early relaxation of coastal upwelling. However, the landings of this small pelagic finfish are also highly variable from year to year.  Small pelagics often exhibit high variability due to the effects of environmental conditions on survival and recruitment [@Checkleyetal2009].  In addition, the Indian oil sardine fishery is largely coastal and artisanal thus migration of fish schools in and out of the coastal zone greatly affects exposure to the fishery.

Researchers have examined a variety of environmental variables for their correlation with landings of the Indian oil sardine.  Precipitation during the southwest monsoon [@MurtyEdelman1966; @AntonyRaja1969; @AntonyRaja1972; @Jayaprakash2002] and the day of the monsoon arrival [@Jayaprakash2002] have been widely examined and correlations found, but the reported effects are positive in some studies and negative in others. Precipitation is correlated with monsoon winds which trigger upwelling but may also affect salinity and act as a direct or indirect cue for spawning.  Researchers have also found correlations with various metrics of upwelling intensity during the post-monsoon period [@LonghurstWooster1990; @Madhupratapetal1994; @Jayaprakash2002; @Krishnakumaretal2008; @Thara2011] and with direct measures of productivity, such as nearshore zooplankton and phytoplankton abundance [@Hornell1910; @Nair1952; @NairSubrahmanyan1955; @Madhupratapetal1994; @Georgeetal2012].  Researchers have also found correlations with sea surface temperature (SST); SST  can affect both somatic growth rates and juvenile survival but also can cause fish to move off-shore and away from the shore-based fishery [@Annigeri1969; @PrabhuDhulkhed1970; @Pillai1991].

<!-- objective of the paper -->
In this paper, we revisit the analysis of environmental covariates that correlate with oil sardine catch using data from 1983 to 2015 on seasonal landings and environmental conditions from remote-sensing.  Based on biological information concerning how environmental conditions in the coastal waters affect spawning density, exposure of sardines to the coastal fishery, and larval and juvenile sardine survival, we develop a set of hypotheses about which remote-sensing covariates should correlate with current or future landings.  We use generalized additive modelling of the landing data with remote-sensing covariates to test for these correlations.  

## Study area and oil sardine life history

<!-- study area -->
Our analysis focuses on the Kerala coast (Figure `r ref("fig:studyarea")`) region of India, where the majority of the Indian oil sardines are landed and where oil sardines comprise ca. 40% of the marine fish catch [@Srinath1998; @Vivekanandanetal2003]. This area is in the Southeast Arabian Sea (SEAS), one of world’s major upwelling zones, with seasonal peaks in primary productivity driven by upwelling caused by winds during the Indian summer monsoon [@Madhupratapetal2001; @Habeebrehmanetal2008] between June and September. Within the SEAS, the coastal zone off Kerala between 9$^\circ$N to 13$^\circ$N has especially intense upwelling due to the combined effects of wind stress and remote forcing [@Smitha2010; @Smithaetal2008].  The result is a strong temperature differential between the near-shore and off-shore and high primary productivty and chlorophyll in this region [@Madhupratapetal2001;  @Habeebrehmanetal2008; @Jayarametal2010; @Smitha2010; @Raghavanetal2010; @Chauhanetal2011].  The primary productivity peaks subside after September while mesozooplankton abundances increase and remain high in the inter-monsoon period [@Madhupratapetal2001].

<!-- life cycle of the oil sardine -->
The life cycle  of the oil sardine fishery begins with the entry of adult fish into the inshore areas during June – July, corresponding with the onset of the southwest monsoon [@Chidambaram1950; @AntonyRaja1969] and restricted along the narrow strip of the western India continental shelf, within 20 km from the shore. The fish migrate from offshore to coastal waters and vice versa coinciding with the prevailing wind conditions [@Hornell1910]. A gradual increase in temperature ranging from 26 to 28°C is favorable for their inshore migration [@Chidambaram1950]. According to Nair et al. [-@Nairetal2016], the exact location of oil sardine spawning grounds along the Indian coastline is still unclear; however, it is generally believed that the spawning of oil sardine occurs during the southwest monsoon period (Jun-Jul) when temperature, salinity and suitable food availability are conducive for larval survival [@MurtyEdelman1966; @JayaprakashPillai2000; @Krishnakumaretal2008], although some studies suggest that spawning can occur as late as  September [@AntonyRaja1969; @Hornell1910; @HornellNayudu1923; @PrabhuDhulkhed1970].

After eggs are spawned, they develop rapidly into larvae within 24 hrs [@Nair1959].  The phytoplankton bloom that provides for sardine larvae food is dependent upon nutrient influx from coastal upwelling and runoff from rivers. The blooms start in the south in June, increase in intensity and spread northward up the coast [@Smitha2010]. Oil sardines grow rapidly during their first few months and can reach 40mm with 2 months (reference). Variation in the bloom initiation time and intensity lead to changes in the food supply to sardine larvae and corresponding changes in their growth rate and survival and the later recruitment of 0-year sardines into the fishery [@Georgeetal2012]. During March to May, the inshore waters warm considerably and sardines move off-shore to deeper waters [@Chidambaram1950].  Catches of sardines are correspondingly low during this time for all size classes.

The age at first maturity occurs at less than one year, at about 150 mm size. Spawning begins during the southwest monsoon period (June-July) when temperature, salinity and suitable food availability are conducive for larval survival and continues into September [@MurtyEdelman1966; @JayaprakashPillai2000;  @Krishnakumaretal2008].  The exact location of oil sardine spawning grounds along the Indian coastline is still unclear [@Nair2016].  The fishery is closed during June to mid-July, and when it commences in mid-July, it is dominated by 1-2.5 year old fish that have recently spawned (spent).  In August a spike of 0-year (40mm) juveniles from the June spawning ofter appears in the catch and another spike of 0-year fish is sometimes seen in February from the last fall spawning.  Overall, catches along the Kerala coast are fairly high throughout the year except during March-May (Figure `r ref("fig:catch")`). 

<!-- What is happening each quarter 
Catch is reported quarterly not monthly. Along the west coast, the length composition of catches ranges between 50-220mm. 1st year spawners (140-160 mm) enter the fishery during June/July and as the season progresses to September-December, adults occur in reduced quantities and catches are dominated by age-1 juveniles (120-140 mm) that were spawned in the prior season (citation). Seasons of rainfall extremes (weak or intense) cause recruitment failure, while a daily rainfall of 20-30 mm during June-August along the southwest coast may indicate good recruitment to the fishery (citation).
-->

<!--
As the success of commercial fishery for each season is determined by the number of juveniles recruited at the beginning of the same season, rainfall, which affects spawning success, has been used to forecast the strength of juvenile brood entering the fishery. Though earlier studies have shown that most of spawning activity occurs in the near-shore area or in the mouths of estuaries, there are observations of eggs and larvae in the shelf region even at 200m depth (citation). Also, studies from other upwelling ecosystems (citation), have shown that the planktonic stage larvae accumulate in frontal zones as well as in river plumes. Studies from Bakun (citation), Bakun and English (citation), Cury and Roy (citation) etc. explain the linkage to the atmospheric and oceanic processes based on number of hypothesis like Optimal environment window, Match-mismatch hypothesis etc. Also, the changes in the fishery in either short term or long term is explained for other similar theories such as population expansion (Mac Call's hypothesis) and the Wasp waist hypothesis.
-->

## Hypotheses




<!-- SST is known to be important in Pacific sardine -->
[@JacobsonMacCall1995; @Checkleyetal2009; @McClatchieetal2010; @LindegrenCheckley2012; @Checkleyetal2017]

<!--
Remote sensing satellites can be used to detect changes in ocean physical, biological and chemical properties, such as surface temperature, winds, surface height, surface waves, rainfall and surface salinity, as well as the ecosystem and water quality changes. These tools are relevant for fisheries management (IOCCG 2009; citation?) as well as development of predictive capabilities.  Add more discussion regarding prior work/references on using satellite data to predict oil sardine catch.  Satellite chlorophyll, SST and SSH have been used to estimate sardine habitat of the west coast of the USA [@Zwolinskietal2011]. In the last few years, satellite measurements of chlorophyll have been used to predict high densities of oil sardines.  This is used a functional tool for producing forecasts for fisherman and has lead to increases in their efficiency (catch per unit effort) (citation)
-->


Based on biological information concerning how environmental conditions affect sardine survival and recruitment and affect exposure of sardines to the coastal fishery, we develop a set of hypotheses about which remote-sensing covariates in which months should correlate with landings in specific quarters. 

### Quarter 3 July-Sept catch correlates

0-yr fish appear in catch.

Quarter 3 is during the summer monsoon, with the peak in June and continuing at a lower level into October.  During this time, sardines move inshore from offshore and become exposed to the primarily coastal fishery.   Factors that affect inshore migration or spawning should affect the catch in quarter 3, which is primarily of mature fish.  The following are the major factors that are hypothesized to affect the catch in quarter 3 due to either a stronger spawning class or higher (or lower) exposure of the spawning age fish to the fishery due to higher (or lower) inshore migration.


* Precipitation low or high during either the summer monsoon (June-July) or pre-monsoon (Apr-May) is associated with low spawning.

* Low sea surface temperature is associated with delayed and limited spawning in Pacific sardines [@JacobsonMacCall1995].  The reasons for this may be behavioral as low temperature as associated with poor survival of larvae.  Larval mortality rates may be higher in colder water due to increased zooplankton predation and upwelling (associated with cold surface temperatures) advects larvae into offshore waters.

* In addition to advection of larvae offshore, upwelling brings low oxygen water to the surface which discourages inshore migration to coastal spawning areas.  Thus due to both lower SST and lower dissolved oxygen, strong upwelling during the spawning season is expected to be negatively associated with the spawning biomass in the near-shore region.

* Salinity is also associated with spawning initiation; however we do not have long-term salinity data.

* Spawners are age 2+ fish, thus the biomass in the previous two seasons should be correlated with the biomass of spawners this year.  This should especially be the case for the biomass in quarters 4, 1 and 2 of the previous season as these were age 1-2 fish.  A stronger cohort of age 1-2 fish in the previous year should be associated with a stronger spawner class in the current year.  Note that it could also be the case that higher catch of age 1-2 fish in the previous season would be associated with lower spawners in the current year due to depletion.  However during the majority of our study period, the fishery was largely artisanal and is not thought to be sufficient to deplete the population.

After spawning, during the 4th quarter of the current calendar year and the 1st and 2nd quarters of the next calendar year, the length distribution of the catch shifts to smaller fish as juveniles begin shoaling inshore (citation).   The catch during the non-spawning part of the season should be correlated with factors associated with increased juvenile shoaling and survival of eggs and larvae from previous seasons’ spawning.  The former is associated with higher exposure of the non-spawning biomass to the coastal fishery and the latter is associated with a stronger (larger) non-spawning biomass.

### Post-summer monsoon catch correlates

The post-spawning period corresponds to catch in quarter 4 of the current calendar year and quarters 1 and 2 of the next calendar year.  The main environmental covariates that should be important are those associated with egg and larval survival as these have been found to be the most important drivers of recruitment (citation) and covariates associated with shoaling and inshore migration of juvenile fish.

* Sea surface temperature is associated with egg and larval survival [@JacobsonMacCall1995].

* Upwelling is associated with higher productivity and higher density of zooplankton, which leads to better larval and juvenile growth and survival.  Thus the strength of upwelling during the post-spawning quarters should be associated with higher non-spawning biomass in subsequent years.  

* Chlorophyll blooms are signatures of high productivity from nutrient influx either due to upwelling or coastal inputs.  Thus the bloom intensity in prior years should be associated with future non-spawning biomass.  In addition, juveniles shoal in response to chlorophyll blooms (citation) and thus the near shore chlorophyll bloom intensity in the current season should be associated with higher post-spawning catches due to higher exposure of the non-spawning biomass to the coastal fishery.

* Fish age 1-2 should be correlated with spawners in the previous two seasons, since a larger spawner class should be associated with higher recruitment.  Thus catch in quarter 3 the previous two calendar years is expected to be correlated with the catch after the summer monsoon (quarter 4 of the current calendar year with quarter 1 and 2 of the next calendar year).

* Because age 2 fish also appear in the catch after quarter 3, we also expect the post-spawning catch in the previous season to be correlated with the post-spawning catch in the current season.

# Materials and methods

## Sardine landing data

Quarterly fish landing data have been collected by the Central Marine Fisheries Research Institute (CMFRI) in Kochi, India, since the early 1950s using a stratified multi-stage sample design that takes into account landing centers, number of fishing days, and boat net combinations in fishing operations. These data are used to produce the quarterly total landings estimates with a precision of ±5% (*CMFRI, 1969-1994; Bal and Rao, 1990, Srinath, 2005*). The quarterly landing data were log-transformed to stabilize the seasonal variance. 

*Add discussion of the composition of the fishery since the early 1950s.   Fishery largely artisanal until recently.  Add discussion of why researchers consider the catch to be an index of total biomass and that fishery is not having a large impact on the standing biomass (because fishery has been traditionally coastal and non-motorized).*

## Covariate data

We analysed monthly composites of the following environmental data derived from satellites: Sea Surface Temperature (SST), chlorophyll-a (CHL), upwelling (UPW) and precipitation. Sea surface temperature: For 1981 to 2002, we used Pathfinder Version 5.2 product on a 4km grid, and for 2003 to 2016, we used Advanced Very-High Resolution Radiometer (AVHRR) data at a 0.1 degree spatial scale.  Chlorophyll-a, we used SeaWiFS data on a 0.1 degree spatial scale for 1981 to 2002 and MODIS data from 2003 to 2017 at 0.05° spatial scale.  The SST and CHL data were averaged across thirteen 1 degree by 1 degree boxes which roughly parallel the bathymetry (Figure `r ref("fig:studyarea")`).  The SST and CHL satellite data were retrieved from the NOAA ERDDAP server (See the Appendices for full details).  

For an index of coastal upwelling, we used the sea-surface temperature differential between near shore and 3 degrees offshore as described by Naidu et al. [-@Naiduetal1999] and Smitha et al. [-@Smithaetal2008].  The index is computed as the average SST in boxes 3 degrees offshore of boxes 1-5 in Figure `r ref("fig:studyarea")` minus the average SST in boxes 1-5.  For SST, we used the remote-sensing sea-surface temperature data sets described above. This SST-based upwelling index has been validated as a more reliable metric of upwelling off the coast of Kerala compared to wind-based upwelling indices [@Smithaetal2008].  

The precipitation data were downloaded from two different sources.  The first was an estimate of the monthly precipitation over Kerala from land-based gauges.  This time series is available from the start of our landing data (1956).  The second was a remote-sensing precipitation product from the NOAA Global Precipitation Climatology Project.  This provides a precipitation estimate using a global 2.5 degree grid.  We used the 2.5 by 2.5 degree box defined by latitude 8.75 to 11.25 and longitude 73.25 to 75.75 for the precipitation off the coast of Kerala.  These data are available from 1979 forward.  Full details on the precipitation data sets are available in the Appendices.  

# Explanatory models

We investigated correlations between environmental variables and sardine catch during spawning and post-spawning periods using generalized additive models [GAMs, @Wood2017].  GAMs allow one to model the effect of a covariate as a flexible non-linear function and it was known that the effects of the environmental covariates were likely to be non-linear, albeit in an unknown way.  Our approach is similar to that taken by Jacobson and MacCall [-@JacobsonMacCall1995] in a study of the effects of SST on Pacific sardine recruitment.   
The model for catch in quarter 3 was of the form
$$ln(S_t) = \alpha + f_1(c_1) + f_2(c_2) + s_1(ln(C_{t-1})) + s_2(ln(C_{t-2})) + \epsilon$$
where $ln(S_t)$ is the log catch in the 3rd quarter of year $t$.  Catch in the 3rd quarter (July-Aug) captures mainly spawning age fish as it overlaps with the tail end of the spawning season---the fishery is closed July to mid-August during the height of the summer monsoon and spawning season.  The model seeks to explain the variability in catch in the 3rd quarter via a non-linear function of the covariates, $c_1$ and $c_2$, plus a non-linear function of total catch in prior years.   The catches were logged to stabilize and normalize the variance.  The model is primarily statistical, meaning it should not be thought of as being a population growth model.  It is a form of model that is often used for modelling the effects of covariates on the number of fish that recruit into the fishery [e.g. @JacobsonMacCall1995, add others or a textbook].  The covariates tested are those discussed in the section on covariates that have been hypothesized to drive the size of the spawning biomass exposed to the fishery.  The relevant covariate may be in the concurrent with the catch or in a prior year.

The model for catch in the non-spawning quarters (October to June) took a similar form
$$ln(N_t) = \alpha + f_1(c_1) + f_2(c_2) + s_1(ln(S_{t-1})) + s_2(ln(N_{t-1})) + \epsilon$$
where $ln(N_t)$ is the log of the total catch in the post-spawning season October to June (a nine-month period that spans two calendar years).   We included the catches during the spawning season ($S_t$) separately from the non-spawning season ($N_t$).  The covariates that affect the egg and larval survival may be quite different from those affecting the age-1 to age-2 or age-2 to age-3 survival.  We wanted to be able to model these separately.  The covariates tested are those discussed in the section on covariates that have been hypothesized to drive both survival and growth of juvenile fish and the factors that lead migration of fish inshore and thus exposed to the coastal fishery.

Model selection was conducted in a step-wise fashion.  The length of the covariate data was 33-35 years (1982 to 2015 or 2017) for most covariates and this was not sufficient for fitting all covariates simultaneously.  The model for the effect of the past of biomass on current biomass (density-dependence) was first determined.  Once a sufficient model for the density-dependence was determined, the covariates were studied individually and then jointly.  F-tests and AIC on nested sets of GAM models [@Woodetal2016] were used to evaluate the support for models.  One feature of GAMs is that they allow the ‘flexibility’ or smoothing parameter of the response curve to be estimated. However we fixed the degree of flexibility so that reasonably smooth responses were achieved.  Multi-modal or overly wiggly response curves would not make sense for our covariates.  We used GAMs with smooth terms represented by penalized regression splines [@Wood2011, using the mgcv package in R] and fixed the smoothing term at an intermediate value.

# Results

## Catches in prior seasons as explanatory variables

There was support for including the post-spawning catch in the previous year as an explanatory variable for the catch during the spawning season (3rd quarter).  Models with $ln(N_{t-1})$ were strongly supported over an intercept only model (Table `r ref("tab:spawner_model_selection")`, time-dependency test).  However the addition of the catch two years prior, $ln(N_{t-2})$, lead to either no decrease in the residual deviance (i.e. increase in the explained variance) and in fact, increased the residual deviance for the model with non-linearity (Table `r ref("tab:spawner_model_selection")`, Linearity test).  We also tested the support for non-linearity in the effect of the prior year catch on the catch in the spawning season.  This was done by comparing models with $ln(N_{t-1})$ included as a linear term or as a non-linear function $s(ln(N_{t-1}))$ (Table `r ref("tab:spawner_model_selection")`, Linearity test).  The residual deviance decreased using a non-linear response however the cost was 1.4 degrees of freedom.  The result was only weak (non-significant) support for allowing a non-linear response.  The full set of models tested, including tests using catch during the spawning months in previous seasons as a covariate are shown in Tables `r ref("tabappA:spawner-model-linear", pre="A")` and `r ref("tabappA:spawner-model-nonlinear", pre="A")`.  The results were the same if we used the full landings data set from 1956 to 2015 (Table `r ref("tabappA:spawner-model-1956", pre="A")`).  Overall, the landings in prior seasons was only weakly explanatory for the catch in the spawning months, and the maximum $R^2$ was less than 30% (Table `r ref("tab:spawner_model_selection")`).

The results were similar for models of the landings  during the non-spawning months ($N_t$) of the sardine season (Table `r ref("tab:nonspawner-models")`).  The most supported model for $N_t$ used a non-linear response to landings during the non-spawning months of the previous season: $ln(N_t) \sim ln(N_{t-1})$ with a non-linear response to landings during the spawning months two seasons prior (Table `r ref("tab:nonspawner-models")`).  There was low support for including landings from the non-spawner months two seasons prior or for using the landings during the spawning months in the prior season (Tables `r ref("tabappA:nonspawner-model-time-a", pre="A")`, `r ref("tabappA:nonspawner-model-time-b", pre="A")`, and `r ref("tabappA:nonspawner-model-time-c", pre="A")`). 

## Environmental covariates as explanatory variables

There was no support for using precipitation during the summer monsoon (Jun-Jul) or pre-monsoon period (Apr-May) as an explanatory variable for the catch during the spawning months (Table `r ref("tabappB:spawner-model-precip", pre="B")`) nor the non-spawning months (Table `r ref("tabappB:nonspawner-model-precip", pre="B")`).  This was the case whether precipitation in the current or previous season was used, if precipitation was included as non-linear or non-linear effect, or if the smoothing term (degree of non-linearity allowed) was estimated and thus not constrained, and if either precipitation during monsoon (Jun-Jul) or pre-monsoon (Apr-May) were used as the covariate.

However, we found significant correlation between average sea surface temperature during the early post-spawning period (Jul-Dec) and catch during the spawning season (Table 3, Table `r ref("tabappB:spawner-model-sst", pre="B")`).  Sea surface temperature has been found to be correlated with sardine biomass in a number of other studies [@JacobsonMacCall1995; add the others].  The residual deviance was lowest in a model with SST in both the prior year and two years prior included and with a non-linear response for both. This a similar result to Jacobson and MacCall [-@JacobsonMacCall1995] who found that SST in multiple prior years was supported as explanatory variables for sardine recruitment and productivity.  However the reduction in degrees of freedom was high for this model and it was not supported, despite having the lowest residual degrees of freedom, given the cost (loss of degrees of freedom).  The model with the lowest AIC was a model with only the current year and a non-linear response.  The response shows a step-response with a negative effect at low temperatures and then an increased effect a higher temperatures (Figure `r ref("fig:cov-effects")`).  This type of step-response has been found in studies of the effect of SST on recruitment in Pacific sardines (*citations from Jacobson and MacCall*).  The $R^2$ for this model was 0.40 (Table `r ref("tab:covariate-models")`).

The strongest predictor of the catch during the spawning season however was the upwelling strength during Jan-Mar (4-6 months prior) in the previous season (Tables `r ref("tab:covariate-models")` and `r ref("tabappB:spawner-model-upw", pre="B")`).  Jan-Mar is a period when the young of the year and age-1 fish would be found feeding in large shoals in the coastal region.  This is also the time of year with the second highest catches and catches that are dominated by small-sized fish (*citation*).  The best supported model included only the upwelling strength in the prior year, without SST included as a covariate (Table `r ref("tab:covariate-models")`).  The $R^2$ for this model was 0.60 and the fitted versus observed catches for this model are shown in Figure `r ref("fig:fittedmodel")`.

For catch in the non-spawning months (Oct-Jun), sea surface temperature during the spawning months in the current season was a significant predictor similar to what was found for catches in the spawning months.  Upwelling in the prior season was also a significant predictor, but the important months were Oct-Dec.  This is period is important for larval and early juvenile survival and growth, and other studies have also found this to be a critical period for future stock size in sardines (*citation*).  The model with the lowest residual variance and lowest AIC was the model which included both the SST and upwelling covariates (Table `r ref("tab:covariate-models")`).  The $R^2$ for this model was 0.73 and the fitted versus observed catches for this model are shown in Figure `r ref("fig:fittedmodel")`.

Chlorophyll-a density is speculated to be an important predictor of larval sardine survival  and growth.  In addition, sardines shoal in response to coastal chlorophyll blooms, which brings them in contact with the coastal fisheries.  Thus Cholorophyll-a density is assumed to be an important either covariate or driver of future or current sardine catches.  However, we had chlorophyll-a remote-sensing data only from 1998 onward. Our simplest covariate model required 5 degrees of freedom, thus we were limited in the analyses we could conduct.  In addition, the years, 1998-2014, have relatively low variability in catch sizes; the logged catch sizes during this period range from 10-11 during the spawning months and 11-12 during the nonspawning months.  Second degree polynomial models were fit (Appendix C) to the average log chlorophyll-a density in the current and prior season from quarters 3 (Jul-Sep), 4 (Oct-Dec), and 1 (Jan-Mar).  Chlorophyll-a density was not a significant predictor for the spawning catch for any of the tested combinations of current or prior season and quarter.  The only significant effect was seen for non-spawner catches using chlorophyll-a density in Oct-Dec of the current and prior season (Table `r ref("tabappC:nonspawner-model", pre="C")`).  This matches results which found that the upwelling index in Oct-Dec of the prior season was a predictor for the non-spawning months' catch.  The upwelling index and chlorophyll-a density are both indices of low-trophic level productivity.

# Discussion

Needs to be written.  The results make sense biologically and are supported by other studies.

Discuss why precipitation did not come up.
Implications.  Supports the use of SST and upwelling indices for forecasting.

# References

<div id="refs"></div>

\clearpage

# Figure Legends

Figure `r ref("fig:studyarea")`. Close up of Kerala State with the latitude/longitude boxes used for the satellite data. Kerala State is marked in grey and the oil sardine catch from this region is being modeled.

```{r catch-dat, echo=FALSE,message=FALSE,warning=FALSE}
dat=oilsardine_qtr
catchmeans=round(tapply(dat$Kerala, dat$Qtr, mean, na.rm=TRUE))
names(catchmeans)=paste("Q",1:4,sep="")
catchmeds=round(tapply(dat$Kerala, dat$Qtr, median, na.rm=TRUE))
names(catchmeds)=paste("Q",1:4,sep="")
```

Figure `r ref("fig:catch")`. Quarterly catch data 1956-2014 from Kerala. The catches have a strong seasonal pattern with the highest catches in quarter 4  Note that quarter 3 is July-Sept and that the fishery is closed July 1 to Aug 15, thus the fishery is only open 1.5 months in quarter 3. The mean catch (metric tonnes) in quarters 1 to 4 are `r catchmeans["Q1"]`, `r catchmeans["Q2"]`, `r catchmeans["Q3"]`, and `r catchmeans["Q4"]`, respectively.

Figure `r ref("fig:cartoon")`.  Cartoon of the sardine life-cycle in the SE Indian Ocean and how it interacts with the fishery. 

Figure `r ref("fig:COVts")`.  Remote sensing covariates used in the analysis.  All data are monthly averages over Box 4 in Figure `r ref("fig:studyarea")` on the Kerala coast off of Kochi.  Panel A) Upwelling Index. The upwelling index is the difference between the near-shore sea surface temperature (SST) and the off-shore SST defined as 3 degrees longitude offshore. Panel B) Surface chlorophyll-a (Chl-a).  The Chl-a data are only available from 1997 onward.  Panel C) Sea surface temperature constructed from Advanced Very High Resolution Radiometer (AVHRR). Panel D) Average daily rainfall (mm/day) off the Kerala coast.

Figure `r ref("fig:SSTts")`.  Key oil sardine life-history events overlaid on the monthly SST in the near-shore and off-shore and the near-shore Chl-a.

Figure `r ref("fig:cov-effects")`.  Effects of covariates estimated from the GAM models. Panel A) Effect of SST during the spawning months (Jul-Sep) on catch during the spawning months.  Low SST is associated with lower that expected catch during the spawning months.  Panel B) Effect of upwelling (inshore/off-shore SST differential) during Jan-Mar of the prior season on catch during the next spawning months (Jul-Sep).  The index is the difference between offshore and inshore SST, thus a negative value indicates warmer coastal surface water than off-shore.  Warm coastal water during Jan-Mar when sardines are foraging along the coast, is associated with lower catch during the next spawning season.  Panel C) Effect of SST during the spawning months (Jul-Sep) on catch during the subsequent non-spawning months (Oct-Jun).  Low SST is associated with lower that expected catch during the following non-spawning months.  Panel D) Effect of upwelling (inshore/off-shore SST differential) during Oct-Dec of the prior season on catch during the non-spawning months (Oct-Jun) the next season.  Strong upwelling (positive upwelling index) in the early larval and juvenile period (Oct-Dec) is associated with higher than expected catch in the next season.

Figure `r ref("fig:fitted")`.  Fitted versus observed catch with models with and without environmental covariates.  Panel A) Fitted versus observed log catch in the spawning months with only non-spawning catch in the previous season as the covariate: $S_t = s(N_{t-1}) + \epsilon$. Panel B) Fitted versus observed log catch in the spawning months with Jan-Mar upwelling in the prior season added as a covariate to the model: $s(V_{t-1})$. Panel C) Fitted versus observed log catch in the non-spawning months with only non-spawning catch in the previous season and spawning catch two season prior as the covariates: $N_t = s(N_{t-1}) + s(S_{t-2}) + \epsilon$. Panel D) Fitted versus observed log catch in the non-spawning months with current season SST in Jul-Sep ($V$) and Oct-Dec upwelling in the prior season ($W$) added as covariates: $s(V_t) + \beta W_{t-1}$.  $W$ was added linearly since the data were insufficient to estimate four non-linear effects.

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = 'Tables.Rmd', echo=FALSE}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = 'Figures.Rmd', echo=FALSE}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# Appendices

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = 'Appendices.Rmd', echo=FALSE}
```

