---
title: Influence of temperature and upwelling intensity on oil sardine (*Sardinella longiceps*) landings along the southwest coast of India
author: ''
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document: default
  md_document: default
  word_document: default
  html_document: default
header-includes:
- \usepackage[labelformat=empty]{caption}
- \usepackage{setspace}
- \onehalfspacing
- \usepackage{etoolbox}
- \AtBeginEnvironment{tabular}{\singlespacing}
- \setlength\parindent{24pt}
- \usepackage{mathptmx}
- \usepackage[margin=1in]{geometry}
- \usepackage[fontsize=12.5pt]{scrextend}
csl: '`r system.file("docs", "ecology.csl", package = "SardineForecast")`'
bibliography: '`r system.file("docs", "Sardine.bib", package = "SardineForecast")`'
---

```{r child = 'setup.Rmd'}
#Generic setup code common across all Rmds
```

# Introduction

<!-- indian oil sardine is important -->
The Indian oil sardine (*Sardinella longiceps* Valenciennes, 1847) is one of the most commercially important fish resources along the southwest coast of India (Figure `r ref("fig:studyarea")`) and historically has comprised approximately 25% of the catch biomass [@Vivekanandanetal2003].  Landings of the Indian oil sardine are highly seasonal and peak during and after the summer monsoon period (June through September), in conjunction with the onset and early relaxation of coastal upwelling. However, the landings of this small pelagic finfish are also highly variable from year to year.  Small pelagics often exhibit high variability due to the effects of environmental conditions on survival and recruitment [@Checkleyetal2009].  In addition, the Indian oil sardine fishery is largely coastal and artisanal thus migration of fish schools in and out of the coastal zone greatly affects exposure to the fishery.

<!-- review of looking at env vars -->
Researchers have examined a variety of environmental variables for their correlation with landings of the Indian oil sardine.  Precipitation during the southwest monsoon [@MurtyEdelman1966; @AntonyRaja1969; @AntonyRaja1974; @Jayaprakash2002] and the day of the monsoon arrival [@Jayaprakash2002] is thought to act as either a direct or indirect cue for spawning.  Many studies have looked for correlations between precipitation, however the reported effects are positive in some studies and negative in others. Researchers have also found correlations with various metrics of upwelling intensity during the post-monsoon period [@LonghurstWooster1990; @Madhupratapetal1994; @Jayaprakash2002; @Krishnakumaretal2008; @Thara2011] and with direct measures of productivity, such as nearshore zooplankton and phytoplankton abundance [@Hornell1910; @Nair1952; @NairSubrahmanyan1955; @Madhupratapetal1994; @Georgeetal2012].  Researchers have also found correlations with sea surface temperature (SST); SST  can affect both somatic growth rates and juvenile survival but also can cause fish to move off-shore and away from the shore-based fishery [@Annigeri1969; @PrabhuDhulkhed1970; @Pillai1991].


In this paper, we revisit forecasting of oil sardine landings using environmental covariates to improve forecast accuracy.  We specifically focus on environmental covariates measured via remote-sensing.  Remote-sensing data give us time series of environmental data over a wide spatial extent at a daily and monthly resolution and at low cost.  A better understanding of how and whether remote sensing data explains variation in seasonal catch will give insights into how the environment affects landings and how (and whether) remote sensing can be used to help improve our forecasts. To model landings, we use autoregressive time-series models (Box-Jenkins models).  Forecasting landings data (catch) using time-series models has a long tradition in fisheries and has been applied to many species [@Mendelssohn1981; @CohenStone1987; @NobelSathianandan1991; @StergiouChristou1996;  @Lloretetal2000; @Georgakarakosetal2006; @Pristaetal2011; @Lawer2016 ], including oil sardines [@Srinath1998; @VenugopalanSrinath1998].  Unlike prior work on landings models with covariates, we use non-linear time-series models to allow a flexible effect of covariates and past catch on current landings.


<!-- Based on biological information concerning how environmental conditions in the coastal waters affect spawning density, exposure of sardines to the coastal fishery, and larval and juvenile sardine survival, we develop a set of hypotheses about which remote-sensing covariates should correlate with current or future landings.  We use generalized additive modelling of the landing data with remote-sensing covariates to test for these correlations.  -->
<!-- closing is objective of the paper -->

<!-- ## Study area and oil sardine life history -->
Our analysis focuses on the Kerala coast (Figure `r ref("fig:studyarea")`) region of India, where the majority of the Indian oil sardines are landed and where oil sardines comprise ca. 40% of the marine fish catch [@Srinath1998; @Vivekanandanetal2003]. This area is in the Southeast Arabian Sea (SEAS), one of world’s major upwelling zones, with seasonal peaks in primary productivity driven by upwelling caused by winds during the Indian summer monsoon [@Madhupratapetal2001; @Habeebrehmanetal2008] between June and September. Within the SEAS, the coastal zone off Kerala between 9$^\circ$N to 13$^\circ$N has especially intense upwelling due to the combined effects of wind stress and remote forcing [@Smitha2010; @Smithaetal2008].  The result is a strong temperature differential between the near-shore and off-shore and high primary productivity and chlorophyll in this region [@Madhupratapetal2001;  @Habeebrehmanetal2008; @Jayarametal2010; @Smitha2010; @Raghavanetal2010; @Chauhanetal2011].  The primary productivity peaks subside after September while mesozooplankton abundances increase and remain high in the inter-monsoon period [@Madhupratapetal2001].

<!-- life cycle of the oil sardine -->
The Indian oil sardine fishery is restricted along the narrow strip of the western India continental shelf, within 20 km from the shore.  The yearly cycle (Figure `r ref("fig:cartoon")`) of the fishery begins at the start of spawning during June – July, corresponding with the onset of the southwest monsoon [@Chidambaram1950; @AntonyRaja1969].  The mature fish migrate from offshore to coastal waters coinciding with the prevailing wind conditions [@Hornell1910]. A gradual increase in temperature ranging from 26 to 28°C is favorable for their inshore migration [@Chidambaram1950], and the spawning begins during the southwest monsoon period when temperature, salinity and suitable food availability are conducive for larval survival [@MurtyEdelman1966; @JayaprakashPillai2000; @Krishnakumaretal2008; @Nairetal2016].  Although peak spawning occurs in June to July, spawning continues into September [@AntonyRaja1969; @Hornell1910; @HornellNayudu1923; @PrabhuDhulkhed1970] and early- and late-spawning cohorts are evident in the length distributions of the 0-year fish.  Spawning occurs in shallow waters outside of the traditional range of the fishery [@AntonyRaja1964], and after spawning the adults migrate closer to the coast and the spent fish become exposed to the fishery.

After eggs are spawned, they develop rapidly into larvae within 24 hrs [@Nair1959].  The phytoplankton bloom that provide sardine larvae food is dependent upon nutrient influx from coastal upwelling and runoff from rivers during the monsoon and early fall. The blooms start in the south near the southern tip of India in June, increase in intensity and spread northward up the coast [@Smitha2010]. Variation in the bloom initiation time and intensity leads to changes in the food supply and to corresponding changes in the growth and survival of larvae and in the later recruitment of 0-year sardines into the fishery [@Georgeetal2012].  Oil sardines grow rapidly during their first few months, and 0-year fish from 40mm to 100mm appear in the catch in August and September in most years [@Nairetal2016; @AntonyRaja1970].  As the phytoplankton bloom spreads northward, the oil sardines follow thus the oil sardine fishery builds from south to north during the post-monsoon period. Oil sardines remain inshore feeding throughout the winter months, until March to May when the inshore waters warm considerably and sardines move off-shore to deeper waters [@Chidambaram1950].  Catches of sardines are correspondingly low during this time for all size classes.  The age at first maturity occurs at approximately 150 mm size [@Nairetal2016], which is reached within one year.  When the monsoon returns, the oil sardine cycle begins anew.

Overall, catches along the Kerala coast are fairly high throughout the year except during March-May (Figure `r ref("fig:catch")`). The age-distribution caught by the fishery varies through the year.  The fishery is closed during June to mid-July during the monsoon and peak spawning, and when it resumes in mid-July, it is first dominated by 1-2.5 year old mature fish [@AntonyRaja1969; @Bensam1964a; @Nairetal2016].   In August or September a spike of 0-year (40mm) juveniles from the June spawning typically appears in the catch [@AntonyRaja1969;  @Nairetal2016] and another spike of 0-year fish is sometimes seen in February from the last fall spawning [@PrabhuDhulkhed1967; @PrabhuDhulkhed1970].  From October through July, the catch is dominated by fish from 120mm-180mm [@PrabhuDhulkhed1970; @AntonyRaja1970; @Nairetal2016] which is a mix of 0-yr, 1-yr and 2-yr fish [@Nairetal2016].  

<!-- What is happening each quarter 
Catch is reported quarterly not monthly. Along the west coast, the length composition of catches ranges between 50-220mm. 1st year spawners (140-160 mm) enter the fishery during June/July and as the season progresses to September-December, adults occur in reduced quantities and catches are dominated by age-1 juveniles (120-140 mm) that were spawned in the prior season (citation). Seasons of rainfall extremes (weak or intense) cause recruitment failure, while a daily rainfall of 20-30 mm during June-August along the southwest coast may indicate good recruitment to the fishery (citation).
-->

<!--
As the success of commercial fishery for each season is determined by the number of juveniles recruited at the beginning of the same season, rainfall, which affects spawning success, has been used to forecast the strength of juvenile brood entering the fishery. Though earlier studies have shown that most of spawning activity occurs in the near-shore area or in the mouths of estuaries, there are observations of eggs and larvae in the shelf region even at 200m depth (citation). Also, studies from other upwelling ecosystems (citation), have shown that the planktonic stage larvae accumulate in frontal zones as well as in river plumes. Studies from Bakun (citation), Bakun and English (citation), Cury and Roy (citation) etc. explain the linkage to the atmospheric and oceanic processes based on number of hypothesis like Optimal environment window, Match-mismatch hypothesis etc. Also, the changes in the fishery in either short term or long term is explained for other similar theories such as population expansion (Mac Call's hypothesis) and the Wasp waist hypothesis.
-->


# Materials and Methods

## Sardine landing data

Quarterly fish landing data have been collected by the Central Marine Fisheries Research Institute (CMFRI) in Kochi, India, since the early 1950s using a stratified multi-stage sample design that takes into account landing centers, number of fishing days, and boat net combinations in fishing operations [@Srinathetal2005]. The quarterly landings for oil sardine landed from all gears in Kerala were obtained from CMFRI reports (1956-1984) and online database (1985-2015). [@CMFRI1969; @Pillai1982; @Jacobetal1987; @CMFRI1995; @CMFRI2016]. The quarterly landing data were log-transformed to stabilize the seasonal variance. 

## Remote-sensing data

We analysed monthly composites of the following environmental data derived from satellites: Sea Surface Temperature (SST), chlorophyll-a (CHL), upwelling (UPW) and precipitation. The covariate time series wna monthly means are shown in Figure `r ref("fig:COVts")`.

For sea surface temperature, we used Advanced Very-High Resolution Radiometer (AVHRR) data, which provides nearshore SST values. Although the ICOADS product provides SST values for earlier years, ICOADS does not provide accurate nearshore temperatures. For 1981 to 2003, we used the Pathfinder Version 5.2 product on a 0.0417 degree grid. These data were provided by GHRSST
and the US National Oceanographic Data Center. For 2004 to 2016, we used the NOAA CoastWatch sea SST products derived from NOAA's Polar Operational Environmental Satellites (POES). For Chlorophyll-a, we used the Chlorophyll-a products developed by the Ocean Biology Processing Group in the Ocean Ecology Laboratory at the NASA Goddard Space Flight Center. For 1997 to 2002, we used the Chlorophyll-a 2014.0 Reprocessing (R2014.0) product from the Sea-viewing Wide Field-of-view 
Sensor (SeaWiFS) on the Orbview-2 satellite. These data are on a 0.1 degree grid.  For 2003 to 2017, we used the MODIS-Aqua product on a 0.05 degree grid. These CHL data are taken from measurements 
by the Moderate Resolution Imaging Spectroradiometer (MODIS) on NASA's Aqua Spacecraft.  The SST and CHL data were averaged across thirteen 1 degree by 1 degree boxes which roughly parallel the bathymetry (Figure `r ref("fig:studyarea")`).  The SST and CHL satellite data were retrieved from the NOAA ERDDAP server [@Simons2017].  

For an index of coastal upwelling, we used the sea-surface temperature differential between near shore and 3 degrees offshore as described by Naidu et al. [-@Naiduetal1999] and Smitha et al. [-@Smithaetal2008].  The index was computed as the average SST in box 4 of Kochi (Figure `r ref("fig:studyarea")`) minus the average SST in box 13  For SST, we used the remote-sensing sea-surface temperature data sets described above. This SST-based upwelling index has been validated as a more reliable metric of upwelling off the coast of Kerala compared to wind-based upwelling indices [@Smithaetal2008]. The upwelling index and Chlorophyll-a blooms are strongly correlated (Figure `r ref("fig:SSTts")`).

Precipitation data were obtained from two different sources.  The first was an estimate of the monthly precipitation (in mm) over Kerala from land-based rain gauges [@KothawaleRajeevan2017].  This time series is available from the start of our landing data (1956).  The second was a remote-sensing precipitation product from the NOAA Global Precipitation Climatology Project [@Adleretal2016].  This provides estimate of precipitation over the ocean using a global 2.5 degree grid.  We used the 2.5 by 2.5 degree box defined by latitude 8.75 to 11.25 and longitude 73.25 to 75.75 for the precipitation off the coast of Kerala.  These data are available from 1979 forward [@NCEI2017].


## Hypotheses

Environmental variability is known to be a key driver of population variability in small forage fish [@Bakun1996; @Curyetal2000; @Checkleyetal2017]. Based on biological information concerning how environmental conditions affect sardine survival and recruitment and affect exposure of sardines to the coastal fishery, we developed a set of hypotheses (Table `r ref("tab:hypotheses")`) concerning which remote-sensing covariates in which months should correlate with landings in specific quarters.  The quarter 3 (Jul-Sep) catch overlaps the monsoon and the main spawning months.  This is also the quarter where small 0-year fish from early spawning (June) often appear in the catch, sometimes in large numbers.  Variables that are affect or are correlated with movement of sardines inshore should be correlated with quarter 3 landings.  In addition,  pre-spawning (Jan-May) conditions should be correlated with the spawning strength as adult oil sardines experience an acceleration of growth during this period along with egg development.  The post-monsoon catch (Oct-May) is a mix of 0-yr fish (less than 12 months old) and mature fish (greater than 12 months old).  Variables that are correlated with larval and juvenile survival should correlate with the post-monsoon catch both in the current year and in futures years one to two years after.  

Our hypotheses (Table `r ref("tab:hypotheses")`) focus mainly on two drivers: upwelling and ocean temperature.  We also test hypotheses concerning precipitation as this has historically been an environmental covariate considered to influence the timing of oil sardine landings.  Salinity has been proposed as responsible for spawning initiation; however we do not have long-term salinity data.

<!--

### Upwelling

and small pelagics in upwelling systems such as that off the coast of SW India coast are  particular influenced by upwelling intensity [].  Upwelling drives both the productivity in these systems (sardine food) and temperature with affects survival and somatic growth.  

1. In addition to advection of larvae offshore, upwelling brings low oxygen water to the surface which discourages inshore migration to coastal spawning areas.  Thus due to both lower SST and lower dissolved oxygen, strong upwelling during the spawning season is expected to be negatively associated with the spawning biomass in the near-shore region.

1. Upwelling is associated with higher productivity and higher density of zooplankton, which leads to better larval and juvenile growth and survival.  Thus the strength of upwelling during the post-spawning quarters should be associated with higher non-spawning biomass in subsequent years.  

1. Chlorophyll blooms are signatures of high productivity from nutrient influx either due to upwelling or coastal inputs.  Thus the bloom intensity in prior years should be associated with future non-spawning biomass.  In addition, juveniles shoal in response to chlorophyll blooms (citation) and thus the near shore chlorophyll bloom intensity in the current season should be associated with higher post-spawning catches due to higher exposure of the non-spawning biomass to the coastal fishery.


### Sea Surface Temperature

1. Low sea surface temperature is associated with delayed and limited spawning in Pacific sardines [@JacobsonMacCall1995].  The reasons for this may be behavioral as low temperature as associated with poor survival of larvae.  Larval mortality rates may be higher in colder water due to increased zooplankton predation and upwelling (associated with cold surface temperatures) advects larvae into offshore waters.

The catch during the post-monsoon part of the season should be correlated with factors associated with increased survival of eggs and larvae the current and previous years.

1. Sea surface temperature is associated with egg and larval survival [@JacobsonMacCall1995].


### Precipitation

1. Precipitation low or high during either the summer monsoon (June-July) or pre-monsoon (Apr-May) is associated with low spawning.

### Prior catch

1. Spawners are age 2+ fish, thus the biomass in the previous two seasons should be correlated with the biomass of spawners this year.  This should especially be the case for the biomass in quarters 4, 1 and 2 of the previous season as these were age 1-2 fish.  A stronger cohort of age 1-2 fish in the previous year should be associated with a stronger spawner class in the current year.  Note that it could also be the case that higher catch of age 1-2 fish in the previous season would be associated with lower spawners in the current year due to depletion.  However during the majority of our study period, the fishery was largely artisanal and is not thought to be sufficient to deplete the population.

1 Abundance of 1-year and 2-year fish should be correlated with strength of the cohorts from the previous two seasons.  We do not have have direct measures of recruitment or spawners. However the catch in quarter 3 is dominated by mature fish, thus catch in quarter 3 in the previous two calendar years is expected to be correlated with the post-monsoon catch after the summer monsoon (quarter 4 of the current calendar year with quarter 1 and 2 of the next calendar year).

1. Because age 2 fish also appear in the catch after quarter 3, we also expect the post-monsoon catch in the previous season to be correlated with the post-spawning catch in the current season.

*Add discussion of the composition of the fishery since the early 1950s.   Fishery largely artisanal until recently.  Add discussion of why researchers consider the catch to be an index of total biomass and that fishery is not having a large impact on the standing biomass (because fishery has been traditionally coastal and non-motorized).*

-->


## Explanatory models

We modeled the catches during the late-spawning season (quarter 3) separately from the post-summer monsoon season (October to June).  Thus there is no seasonalily in our catch time series, as we analyzed a yearly time series of quarter 3 catches separately from a yearly time series of post-monsoon catches.  We divided the catch in this way for biological reasons.  Catch in the 3rd quarter (July-Sep) captures a mix of spawning age fish as it overlaps with the tail end of the spawning season, is affected by a fishery closure from July to mid-August during the monsoon, and is periodically inflated by the appearance of small 0-year fish from early summer spawning.  In addition, the covariates that affect the timing of spawning and the early egg and larval survival may be different than those that affect later growth, survival and  shoaling that exposes fish to the inshore fishery.  

We tested ARIMA models on both oil sardine catch time series and found little support for auto-regressive errors (ARIMA models with a MA component) based on diagnostic tests of the residuals and model selection. The best supported ARIMA models were simple AR models ($x_t = b x_{t-1} + \eta$). This lack of strong autocorrelation in residuals has been found in other studies that tested ARIMA models for forecasting oil sardine catch [@StergiouChristou1996].  We thus used AR-only models, however we tested both linear and non-linear models using a form of generalized additive model (GAM).  The landings models were fit using conditional sum of squares (conditioning on the first 1 or 2 landings values in the time series).

We investigated correlations between environmental variables and sardine catch  using generalized additive models [GAMs, @Wood2017].  GAMs allow one to model the effect of a covariate as a flexible non-linear function and it was known that the effects of the environmental covariates were likely to be non-linear, albeit in an unknown way.  Our approach is analogous to that taken by Jacobson and MacCall [-@JacobsonMacCall1995] in a study of the effects of SST on Pacific sardine recruitment.  

The first step in our analysis was to determine the catch model: the model for current catch as a function of the past catch.  One feature of GAMs is that they allow the ‘flexibility’ or smoothing parameter of the response curve to be estimated. However we fixed the degree of flexibility so that reasonably smooth responses were achieved and to limit the flexibility of the models being fit.  Multi-modal or overly flexible response curves would not realistic for our application.  We used GAMs with smooth terms represented by penalized regression splines [@Wood2011, using the mgcv package in R] and fixed the smoothing term at an intermediate value (sp=0.6).

Our catch models took the form

* random walk: $ln(C_{i,t}) = \alpha + ln(C_{j,t-1})) + \epsilon$
* AR-1: $ln(C_{i,t}) = \alpha + \phi_1 ln(C_{j,t-1})) + \epsilon$
* AR-2: $ln(C_{i,t}) = \alpha + \phi_1 ln(C_{j,t-1})) + \phi_2 ln(C_{k,t-2})) + \epsilon$
* non-linear: $ln(C_{i,t}) = \alpha + s_1(ln(C_{j,t-1})) + s_2(ln(C_{k,t-2})) + \epsilon$

where $ln(C_{i,t})$ is the log catch in the current year $t$ in season $i$. We modeled two different seasons: 3rd quarter catch $C_{m,t}$ (Jul-Sep) which is during the late part of the summer monsoon and post-summer monsoon catch $C_{p,t}$ (Oct-Jun).  The catches were logged to stabilize and normalize the variance.  $s()$ is a non-linear function estimated by the GAM algorithm.  The model is primarily statistical, meaning it should not be thought of as being a population growth model.  We tested models with prior year post-summer monsoon catch ($C_{p,t-1}$) and 3rd quarter catch catch ($C_{m,t-1}$) as the explanatory catch variable. The catch models were fit to 1982 to 2015 catch data, corresponding to the years where the SST, upwelling and precipitation data were available.  F-tests and AIC on nested sets of models [@Woodetal2016] were used to evaluate the support for the catch models and later for the covariate models.  After selection the best model with the 1982-2015 data, the fitting was repeated with the 1955 to 1981 catch data to confirm the form of the catch models.

Once the catch models were determined, the covariates were studied individually and then jointly. As with the catch models, F-tests and AIC on nested sets of GAM models [@Woodetal2016] were used to evaluate the support for models with covariates.  The smoothing term was fixed at an intermediate value (sp=0.6) instead of treated as an estimated variable.  Our models for catch with covariates took the form
$$ln(C_{i,t}) = M + f_1(c_1) + f_2(c_2) + \epsilon$$
where $M$ was the best catch model from step 1.  The covariate model seeks to explain unexplained variance in the the catch model via a linear or non-linear function of the covariates, $c_1$ and $c_2$.  The covariates tested are those discussed in the section on covariates that have been hypothesized to drive the size of the biomass exposed to the fishery.  


# Results

## Catches in prior seasons as explanatory variables

There was support for including the post-summer monsoon catch in the previous year as an explanatory variable for the quarter 3 catch (June to August).  Models with $ln(C_{t-1})$ were strongly supported over an intercept only model (Table `r ref("tab:spawner_model_selection")`, time-dependency test).  However the addition of the catch two years prior, $ln(C_{t-2})$, lead to either no decrease in the residual deviance (i.e. increase in the explained variance) and in fact, increased the residual deviance for the model with non-linearity (Table `r ref("tab:spawner_model_selection")`, Linearity test).  We also tested the support for non-linearity in the effect of the prior year catch on the catch in the 3rd quarter.  This was done by comparing models with $ln(C_{t-1})$ included as a linear term or as a non-linear function $s(ln(C_{t-1}))$ (Table `r ref("tab:spawner_model_selection")`, Linearity test).  The residual deviance decreased using a non-linear response however the cost was 1.4 degrees of freedom.  The result was only weak (non-significant) support for allowing a non-linear response.  The full set of models tested, including tests using catch during the spawning months in previous seasons as a covariate are shown in Tables `r ref("tabappA:spawner-model-linear", pre="A")` and `r ref("tabappA:spawner-model-nonlinear", pre="A")`.  The results were the same if we used the full landings data set from 1956 to 2015 (Table `r ref("tabappA:spawner-model-1956", pre="A")`).  Overall, the landings in prior seasons was only weakly explanatory for the catch in the 3rd quarter, and the maximum $R^2$ was less than 30% (Table `r ref("tab:spawner_model_selection")`).

The results were similar for models of the landings ($C_t$) during the post-summer monsoon months (Table `r ref("tab:nonspawner-models")`).  The most supported model for $C_t$ used a non-linear response to landings during the post-monsoon months of the previous season: $ln(C_t) \sim ln(C_{t-1})$ with a non-linear response to landings during the spawning season (3rd quarter) two years prior (Table `r ref("tab:nonspawner-models")`).  There was low support for including landings outside of the spawning months two seasons prior or for using the landings during the spawning months in the immediately prior season (Tables `r ref("tabappA:nonspawner-model-time-a", pre="A")`, `r ref("tabappA:nonspawner-model-time-b", pre="A")`, and `r ref("tabappA:nonspawner-model-time-c", pre="A")`).  We did not test models using the catch during quarter 3 in the current fishing season as these data would not be available; the data requires 12 months to process.

## Environmental covariates as explanatory variables

There was no support for using precipitation during the summer monsoon (Jun-Jul) or pre-monsoon period (Apr-May) as an explanatory variable for the catch during the spawning months (Table `r ref("tabappB:spawner-model-precip", pre="B")`) nor the non-spawning months (Table `r ref("tabappB:nonspawner-model-precip", pre="B")`).  This was the case whether precipitation in the current or previous season was used, if precipitation was included as non-linear or non-linear effect, or if the smoothing term (degree of non-linearity allowed) was estimated and thus not constrained, and if either precipitation during monsoon (Jun-Jul) or pre-monsoon (Apr-May) were used as the covariate.

However, we found significant correlation between average sea surface temperature during the early post-spawning period (Jul-Dec) and catch during the spawning season (Table 3, Table `r ref("tabappB:spawner-model-sst", pre="B")`).  Sea surface temperature has been found to be correlated with sardine biomass in a number of other studies [@JacobsonMacCall1995; add the others].  The residual deviance was lowest in a model with SST in both the prior year and two years prior included and with a non-linear response for both. This a similar result to Jacobson and MacCall [-@JacobsonMacCall1995] who found that SST in multiple prior years was supported as explanatory variables for sardine recruitment and productivity.  However the reduction in degrees of freedom was high for this model and it was not supported, despite having the lowest residual degrees of freedom, given the cost (loss of degrees of freedom).  The model with the lowest AIC was a model with only the current year and a non-linear response.  The response shows a step-response with a negative effect at low temperatures and then an increased effect a higher temperatures (Figure `r ref("fig:cov-effects")`).  This type of step-response has been found in studies of the effect of SST on recruitment in Pacific sardines [@JacobsonMacCall1995].  The $R^2$ for this model was 0.40 (Table `r ref("tab:covariate-models")`).

The strongest predictor of the catch during the spawning season however was the upwelling strength during Jan-Mar (4-6 months prior) in the previous season (Tables `r ref("tab:covariate-models")` and `r ref("tabappB:spawner-model-upw", pre="B")`).  Jan-Mar is a period when the young of the year and age-1 fish would be found feeding in large shoals in the coastal region.  This is also the time of year with the second highest catches and catches that are dominated by small-sized fish (*citation*).  The best supported model included only the upwelling strength in the prior year, without SST included as a covariate (Table `r ref("tab:covariate-models")`).  The $R^2$ for this model was 0.60 and the fitted versus observed catches for this model are shown in Figure `r ref("fig:fittedmodel")`.

For catch in the post-summer monsoon months (Oct-Jun), sea surface temperature during July to September in the current season was a significant predictor similar to what was found for catches in the spawning months (quarter 3).  Upwelling in the prior season was also a significant predictor, but the important months were Oct-Dec.  This is period is important for larval and early juvenile survival and growth, and other studies have also found this to be a critical period for future stock size in sardines (*citation*).  The model with the lowest residual variance and lowest AIC was the model which included both the SST and upwelling covariates (Table `r ref("tab:covariate-models")`).  The $R^2$ for this model was 0.73 and the fitted versus observed catches for this model are shown in Figure `r ref("fig:fittedmodel")`.

Chlorophyll-a density is speculated to be an important predictor of larval sardine survival  and growth.  In addition, sardines shoal in response to coastal chlorophyll blooms, which brings them in contact with the coastal fisheries.  Thus chlorophyll-a density is assumed to be an important driver of future or current sardine catches.  We had chlorophyll-a remote-sensing data only from 1998 onward.  Our simplest covariate model required 5 degrees of freedom, thus we were limited in the analyses we could conduct.  In addition, the years, 1998-2014, have relatively low variability in catch sizes; the logged catch sizes during this period range from 10-11 during quarter 3 and 11-12 during the other three quarters.  Second degree polynomial models were fit (Appendix C) to the average log chlorophyll-a density in the current and prior season from quarter 3 (Jul-Sep), 4 (Oct-Dec), and 1 (Jan-Mar).  Chlorophyll-a density was not a significant predictor for the spawning catch for any of the tested combinations of current or prior season and quarter.  The only significant effect was seen for post-summer monsoon catches using chlorophyll-a density in Oct-Dec of the current and prior season (Table `r ref("tabappC:nonspawner-model", pre="C")`).  This matches results which found that the upwelling index in Oct-Dec of the prior season was a predictor for the post-summer monsoon catch.  The upwelling index and chlorophyll-a density are both indices of low-trophic level productivity.

# Discussion

Sardines in all the world's ecosystems exhibit large fluctuations in abundance [@Schwartzloseetal2010; @Baumgartneretal1992].  These small forage fish are strongly influenced by natural variability in upwelling driven by both large-scale forces, i.e. El Nino patterns, and by changes in winds and currents and in addition local conditions of temperature, salinity, and oxygen levels have both direct and indirect on sardine recruitment and survival.

Many studies on Pacific sardines have looked at the correlation between sea surface temperature (SST) and recruitment.  Temperature can have direct effect, an indirect effect on food availability or affect survival [@Houde1987].  Studies in the California Current System, have found that SST explains year-to-year variability in Pacific sardine recruitment [@JacobsonMacCall1995; @Checkleyetal2009; @LindegrenCheckley2012; @Checkleyetal2017].  Consistent with these studies, we also found that SST was the covariate that explained variability in catch anomalies (difference between the landing prediction from prior years' catches).  

McClatchie et al. [-@McClatchieetal2010] found no SST relationship with SST and Pacific sardine recruitment, however their analysis used a linear relationship while both other studies that found a relationship [@JacobsonMacCall1995; @Checkleyetal2017] allowed a non-linear relationship.  Both Jacobson and MacCall (1995) and Checkley et al (2017) found a step-like response function for temperature, where lower temperatures were poor and had negative effects and then at a threshold value the effect became positive.  Our analysis found a similar step-like effect function (Figure ). 

There were three outlier years when catch were much lower than expected based on prior catches.  For these years, sea surface temperature improved the model fit greatly (Figure `r ref("fig:fitted")`). 

Implications.  In California Current system, SST is used as an indicator for recruitment and is used in management of harvest.

<!--
Remote sensing satellites can be used to detect changes in ocean physical, biological and chemical properties, such as surface temperature, winds, surface height, surface waves, rainfall and surface salinity, as well as the ecosystem and water quality changes. These tools are relevant for fisheries management (IOCCG 2009; citation?) as well as development of predictive capabilities.  Add more discussion regarding prior work/references on using satellite data to predict oil sardine catch.  Satellite chlorophyll, SST and SSH have been used to estimate sardine habitat of the west coast of the USA [@Zwolinskietal2011]. In the last few years, satellite measurements of chlorophyll have been used to predict high densities of oil sardines.  This is used a functional tool for producing forecasts for fisherman and has lead to increases in their efficiency (catch per unit effort) (citation)
-->


# References

<div id="refs"></div>

\clearpage

# Figure Legends

Figure `r ref("fig:studyarea")`. Close up of Kerala State with the latitude/longitude boxes used for the satellite data. Kerala State is marked in grey and the oil sardine catch from this region is being modeled.

```{r catch-dat, echo=FALSE,message=FALSE,warning=FALSE}
dat=oilsardine_qtr
catchmeans=round(tapply(dat$Kerala, dat$Qtr, mean, na.rm=TRUE))
names(catchmeans)=paste("Q",1:4,sep="")
catchmeds=round(tapply(dat$Kerala, dat$Qtr, median, na.rm=TRUE))
names(catchmeds)=paste("Q",1:4,sep="")
```

Figure `r ref("fig:catch")`. Quarterly catch data 1956-2014 from Kerala. The catches have a strong seasonal pattern with the highest catches in quarter 4  Note that quarter 3 is July-Sept and that the fishery is closed July 1 to Aug 15, thus the fishery is only open 1.5 months in quarter 3. The mean catch (metric tonnes) in quarters 1 to 4 are `r round(catchmeans["Q1"]/1000,1)`, `r round(catchmeans["Q2"]/1000,1)`, `r round(catchmeans["Q3"]/1000,1)`, and `r round(catchmeans["Q4"]/1000,1)` metric tonnes respectively.

Figure `r ref("fig:cartoon")`.  The sardine life-cycle in the SE Indian Ocean and how it interacts with the fishery. 

Figure `r ref("fig:COVts")`.  Remote sensing covariates used in the analysis.  All data are monthly averages over Box 4 in Figure `r ref("fig:studyarea")` on the Kerala coast off of Kochi.  Panel A) Upwelling Index. The upwelling index is the difference between the near-shore sea surface temperature (SST) and the off-shore SST defined as 3 degrees longitude offshore. Panel B) Surface chlorophyll-a (Chl-a).  The Chl-a data are only available from 1997 onward.  Panel C) Sea surface temperature constructed from Advanced Very High Resolution Radiometer (AVHRR). Panel D) Average daily rainfall (mm/day) off the Kerala coast.

Figure `r ref("fig:SSTts")`.  Key oil sardine life-history events overlaid on the monthly SST in the near-shore and off-shore and the near-shore Chl-a.

Figure `r ref("fig:cov-effects")`.  Effects of covariates estimated from the GAM models. Panel A) Effect of SST during the spawning months (Jul-Sep) on catch during the spawning months.  Low SST is associated with lower that expected catch during the spawning months.  Panel B) Effect of upwelling (inshore/off-shore SST differential) during Jan-Mar of the prior season on catch during the next spawning months (Jul-Sep).  The index is the difference between offshore and inshore SST, thus a negative value indicates warmer coastal surface water than off-shore.  Warm coastal water during Jan-Mar when sardines are foraging along the coast, is associated with lower catch during the next spawning season.  Panel C) Effect of SST during the spawning months (Jul-Sep) on catch during the subsequent non-spawning months (Oct-Jun).  Low SST is associated with lower that expected catch during the following non-spawning months.  Panel D) Effect of upwelling (inshore/off-shore SST differential) during Oct-Dec of the prior season on catch during the non-spawning months (Oct-Jun) the next season.  Strong upwelling (positive upwelling index) in the early larval and juvenile period (Oct-Dec) is associated with higher than expected catch in the next season.

Figure `r ref("fig:fitted")`.  Fitted versus observed catch with models with and without environmental covariates.  Panel A) Fitted versus observed log catch in the spawning months with only non-spawning catch in the previous season as the covariate: $S_t = s(N_{t-1}) + \epsilon$. Panel B) Fitted versus observed log catch in the spawning months with Jan-Mar upwelling in the prior season added as a covariate to the model in panel A. This model was: $S_t = s(N_{t-1}) + s(V_{t-1}) + \epsilon$. Panel C) Fitted versus observed log catch in the non-spawning months with only non-spawning catch in the previous season and spawning catch two season prior as the covariates: $N_t = s(N_{t-1}) + s(S_{t-2}) + \epsilon$. Panel D) Fitted versus observed log catch in the non-spawning months with current season SST in Jul-Sep ($V$) and Oct-Dec upwelling in the prior season ($W$) added as covariates. This model was $N_t = s(N_{t-1}) + s(S_{t-2}) + s(V_t) + \beta W_{t-1} + \epsilon$.  $W$ was added linearly since the data were insufficient to estimate four non-linear effects.

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = 'Tables.Rmd', echo=FALSE}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# Appendices

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = 'Appendices.Rmd', echo=FALSE}
```

```{r, include=FALSE}
#kableExtra is needed for the hypotheses table but 
#needs to be after all the other tables. Otherwise it adds lines to them.
library(kableExtra)
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = 'Figures.Rmd', echo=FALSE}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```




