---
title: 'Table 3: Covariate tests'
output:
  word_document: default
  html_document: default
  pdf_document: default
header-includes: \usepackage[labelformat=empty]{caption}
---

```{r child = file.path(here::here(),'setup.Rmd')}
```

```{r child = file.path(here::here(),'AppendixFiles','Appendix-Functions.Rmd')}
```

```{r child = file.path(here::here(),'AppendixFiles','Appendix-Variables.Rmd')}
```


```{r tab3-spawncovnames, echo=FALSE}
covnamelist = list(
    `July-September catch`=c(
      "SST.2.5.mon6to9.", 
      "Bakun.UPW.mon6to9.",
      "Precip.Kerala.mon6to7.",
      "SST.2.10.3.yr.runsum.",
      "DMI.3.yr.runsum.")
    )
covlags = list(
  SST.2.5.mon6to9. = 0,
  Precip.Kerala.mon6to7. = 0,
  Bakun.UPW.mon6to9. = 0,
  DMI.3.yr.runsum. = 0,
  SST.2.10.3.yr.runsum. = 0)
allcovnames <- c()
covnames <- unlist(covnamelist)
for(i in covnames) allcovnames <- c(allcovnames, 
   paste0(i, covlags[[i]]))
datcommon <- dat.spawners[, 
  c("Year", "spawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", allcovnames)]
dat <- na.omit(datcommon)
NE=mean(abs(dat$spawners0-dat$spawners1))
```

```{r tab3-spawnerapptabeqns, echo=FALSE}
nullmod <- "$ln(S_t) = ln(S_{t-1}) + \\epsilon_t$"
basemod <- "$ln(S_t) = \\alpha + s(ln(N_{t-1})) + \\epsilon_t$"
resp <- "ln(S_t)"
baseformula <- "spawners0 ~ s(nspawners1, sp=0.6)"
nullformula <- "spawners0 ~ -1 + offset(spawners1)"
```

```{r tab3-spawntable}
seltable <- maketable3(dat, nullmod, basemod, nullformula, baseformula, resp, covnamelist, covlags, type="nonlinear", include.num=FALSE, include.group=FALSE, header="**July-Sept catch**", compact=TRUE)
```

```{r tab3-nonspawncovnames, echo=FALSE}
covnamelist = list(
    `October-May catch - simpler model`=c(
      "Precip.Kerala.mon6to7.",
      "SST.2.10.3.yr.runsum.",
      "DMI.3.yr.runsum.")
    )
covlags = list(
  SST.2.5.mon6to9. = 0,
  Precip.Kerala.mon6to7. = 0,
  Bakun.UPW.mon6to9. = 0,
  DMI.3.yr.runsum. = 0,
  SST.2.10.3.yr.runsum. = 0)
allcovnames <- c()
covnames <- unlist(covnamelist)
for(i in covnames) allcovnames <- c(allcovnames, 
   paste0(i, covlags[[i]]))
datcommon <- dat.nonspawners[, 
  c("Year", "nspawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", allcovnames)]
dat <- na.omit(datcommon)
NE=mean(abs(dat$spawners0-dat$spawners1))
```

```{r tab3-nonspawnerapptabeqns, echo=FALSE}
nullmod <- "$ln(N_t) = ln(N_{t-1}) + \\epsilon_t$"
basemod <- "$ln(N_t) = \\alpha + s(ln(N_{t-1})) + \\epsilon_t$"
resp <- "ln(N_t)"
baseformula <- "nspawners0 ~ s(nspawners1, sp=0.6)"
nullformula <- "nspawners0 ~ -1 + offset(nspawners1)"
```

```{r tab3-nonspawntable}
nonspawnerwoS2table <- maketable3(dat, nullmod, basemod, nullformula, baseformula, resp, covnamelist, covlags, type="nonlinear", include.num=FALSE, include.group=FALSE, header="**October-March catch - simpler model**", compact=TRUE)
seltable <- rbind(seltable, nonspawnerwoS2table)
```

```{r tab3-nonspawncovnames2, echo=FALSE}
covnamelist = list(
    `October-May catch - more complex model`=c(
      "Precip.Kerala.mon6to7.",
      "SST.2.10.3.yr.runsum.",
      "DMI.3.yr.runsum.")
    )
covlags = list(
  SST.2.5.mon6to9. = 0,
  Precip.Kerala.mon6to7. = 0,
  Bakun.UPW.mon6to9. = 0,
  DMI.3.yr.runsum. = 0,
  SST.2.10.3.yr.runsum. = 0)
allcovnames <- c()
covnames <- unlist(covnamelist)
for(i in covnames) allcovnames <- c(allcovnames, 
   paste0(i, covlags[[i]]))
datcommon <- dat.nonspawners[, 
  c("Year", "nspawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", allcovnames)]
dat <- na.omit(datcommon)
NE=mean(abs(dat$spawners0-dat$spawners1))
```

```{r tab3-nonspawnerapptabeqns2, echo=FALSE}
nullmod <- "$ln(N_t) = ln(N_{t-1}) + \\epsilon_t$"
basemod <- "$ln(N_t) = \\alpha + s(ln(N_{t-1})) + s(ln(S_{t-2})) + \\epsilon_t$"
resp <- "ln(N_t)"
baseformula <- "nspawners0 ~ s(nspawners1, sp=0.6) + s(spawners2, sp=0.6)"
nullformula <- "nspawners0 ~ -1 + offset(nspawners1)"
```

```{r tab3-nonspawntable2}
nonspawnertable <- maketable3(dat, nullmod, basemod, nullformula, baseformula, resp, covnamelist, covlags, type="nonlinear", include.num=FALSE, include.group=FALSE, header="**October-March catch - more complex model**", compact=TRUE)
seltable <- rbind(seltable, nonspawnertable)
```

```{r print-table3, echo=FALSE}
seltable = seltable[,c("Model", "Residual.df", "Adj.R2", "RMSE", "AIC", "LOOCV", "LOOMd")]
colnames(seltable) <- c("Model", "Resid. df", "Adj. R2", "RMSE", "AICc", "LOOCV", "LOOMdAE")
if(isLatex) colnames(seltable)=c("Model", "\\shortstack{Resid.\\\\df}", "\\shortstack{Adj.\\\\$R^2$}", "RMSE", "AICc", "\\shortstack{LOOCV\\\\RMSE}", "\\shortstack{LOOCV\\\\MdAE}")

thecap=paste("Top covariates for the monsoon (Jul-Sep) and post-monsoon (Oct-Mar) catch ($S_t$ and $N_t$) models. M is the base models with only prior catch as covariates. To the base models, the  environmental covariates are added. ns-SST is nearshore (0-80km) and r-SST is regional (0-160km). The full set of nested covariate models and tests are given in the appendices.", sep="")
fullcap=paste("Table ", ref("tab:covariate-models"), ". ", thecap, sep="")
kable(seltable, align=paste0("l", paste0(rep("c",dim(seltable)[2]-1),collapse="")), caption=fullcap, escape=FALSE)
```

Notes: The nested F-tests are given in Supporting Information. LOOCV = Leave one out cross-validation. RMSE = root mean square error. MdAE = median absolute error. AICc = Akaike Information Criterion corrected for small sample size. $\dagger$ and $\dagger\dagger$ = AICc greater than 2 and greater than 5 below model M (base catch model).  $\ddagger$, $\ddagger\ddagger$, and ${\ddagger}{\ddagger}{\ddagger}$  = LOOCV RMSE 5\%, 10\% and 20\% below model M, respectively. $t$ indicates current season (Jul-Jun). For covariates that are multiyear, $t$ is the current calendar year.