---
title: 'Appendices: Non-spawner covariate tests'
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes: \usepackage[labelformat=empty]{caption}
---

```{r echo=FALSE}
addchl = TRUE
add1956 = FALSE
```

```{r child = file.path(here::here(),'setup.Rmd')}
```

```{r child = file.path(here::here(),'AppendixFiles','Appendix-Functions.Rmd')}
```

```{r child = file.path(here::here(),'AppendixFiles','Appendix-Variables.Rmd')}
```


```{r nonspawncovnames2, echo=FALSE}
vars <- getnonspawnercovs()
datcommon <- dat.nonspawners[, 
  c("Year", "nspawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", vars$allcovnames)]
dat <- na.omit(datcommon)

NE=mean(abs(dat$nspawners0-dat$nspawners1))
```

```{r nonspawnertablevals2}
nullmod <- "$ln(N_t) = ln(N_{t-1}) + \\epsilon_t$"
basemod <- "$ln(N_t) = \\alpha + s(ln(N_{t-1})) + \\epsilon_t$"
resp <- "ln(N_t)"
baseformula <- "nspawners0 ~ s(nspawners1, sp=0.6)"
nullformula <- "nspawners0 ~ -1 + offset(nspawners1)"
```

```{r nonspawnertable2}
seltable <- makeapptable(dat, nullmod, basemod, nullformula, baseformula, resp, vars$covnamelist, vars$covlags)
```

```{r nonspawnertablechlvals1, eval=addchl}
basemod <- "$ln(N_t) = \\alpha + p(ln(N_{t-1})) + \\epsilon_t$"
baseformula <- "nspawners0 ~ poly(nspawners1, 2, raw=TRUE)"
m3form <- "poly(cov, 2, raw=TRUE)"
```


```{r nonspawnchlnames2, echo=FALSE, eval=addchl}
vars <- getnonspawnerchlcovs()
datcommon <- dat.chl[, 
  c("Year", "nspawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", vars$allcovnames)]
dat <- na.omit(datcommon)
```

```{r nonspawnerchltable2, eval=addchl}
seltable = rbind(seltable, spacetmp)
chlbody <- makeapptable(dat, nullmod, basemod, nullformula, baseformula, resp, vars$covnamelist, vars$covlags, m3form=m3form)
seltable <- rbind(seltable, chlbody)
```

```{r nonspawncovnames1956b, echo=FALSE, eval=add1956}
covnamelist = list(
    `Select covariates available from 1957`=c(
      "Precip.Kerala.mon6to7.", "DMI.mon9to11.", "DMI.3.yr.runsum.")
    )
covlags = list(
  Precip.Kerala.mon6to7. = c(0),
  DMI.mon9to11. = c(1),
  DMI.3.yr.runsum. = 0)
allcovnames <- c()
covnames <- unlist(covnamelist)
for(i in covnames) allcovnames <- c(allcovnames, 
   paste0(i, covlags[[i]]))

datcommon <- dat.nonspawners[, 
  c("Year", "nspawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", allcovnames)]
dat <- na.omit(datcommon)

basemod <- "$ln(N_t) = \\alpha + s(ln(N_{t-1})) + \\epsilon_t$"
baseformula <- "nspawners0 ~ s(nspawners1, sp=0.6)"
```

```{r nonspawnertable1956b, eval=add1956}
seltable = rbind(seltable, spacetmp)
tabbody <- makeapptable(dat, nullmod, basemod, nullformula, baseformula, resp, covnamelist, covlags)
seltable <- rbind(seltable, tabbody)
```


```{r print-apptable2, echo=FALSE}
seltable = seltable[,c("Model", "Residual.df", "Adj.R2", "RMSE", "AIC", "LOOCV", "LOOMd")]
colnames(seltable) <- c("Model", "Resid. df", "Adj. R2", "RMSE", "AICc", "LOOCV", "LOOMdAE")
if(isLatex) colnames(seltable)=c("Model", "\\shortstack{Resid.\\\\df}", "\\shortstack{Adj.\\\\$R^2$}", "RMSE", "AICc", "\\shortstack{LOOCV\\\\RMSE}", "\\shortstack{LOOCV\\\\MdAE}")

thecap=paste("Covariate tests for the October-March catch ($N_t$). M is the base model with prior season October-March catch ($N_{t-1}$) and July-September catch two seasons prior ($S_{t-2}$) as the covariates. To the base model, the  environmental covariates are added. ns-SST is nearshore (0-80km) and r-SST is regional (0-160km) SST. Similarly, ns-Chl is nearshore chlorophyll. The models are nested sets, e.g. 1, 2a, 3a and 1, 2b, 3b.")
fullcap=paste("Table ", ref("apptab:nonspawnercovariatewoS2-models", pre="A"), ". ", thecap, sep="")
kable(seltable, align=paste0("l", paste0(rep("c",dim(seltable)[2]-1),collapse="")), caption=fullcap, escape=FALSE)
```

Notes: LOOCV = Leave one out cross-validation. RMSE = root mean square error. MdAE = median absolute error. AICc = Akaike Information Criterion corrected for small sample size. $\dagger$ and $\dagger\dagger$ = AICc greater than 2 and greater than 5 below model M (base catch model).  $\ddagger$, $\ddagger\ddagger$, and ${\ddagger}{\ddagger}{\ddagger}$  = LOOCV RMSE 5\%, 10\% and 20\% below model M, respectively. $t$ indicates current season (Jul-Jun) and $t-1$ is prior season. Thus a Jan-Mar covariate with $t-1$ would be in the same calendar year as the Jul-Sep catch, though in a prior fishing season. With the exception that for covariates that are calendar year (Jan-Dec) or multiyear, $t$ is the current calendar year.